Error-Centric Annotation of Learner Corpora

### 2009-04-09 text/paper.tex
Final!

Upon inspecting the corpus data in greater detail, we found out that it was
sometimes hard to decide whether a particular error was a contextual or
grammatic error. Moreover, often the categories within the grammatic error
branch were not clear enough. It seems that our merger of the CLC and Free Text
introduced some confusion with certain categories. Particularly, the
\textit{omission}, \textit{replace}, and \textit{redundancy} cases would at
times `steal' hits from the other categories due to their very general
nomenclature. We have revised the annotation manual to specifically mention that
all other grammatical error categories take precedence over these three.
Unfortunately, this seems to be a suboptimal solution. The next section contains
a proposal for a much cleaner solution.

@@

Another very unclear point in the original manual was the minimalization of
error spans constraint, which advises the annotator to only select the smallest
possible amount of tokens, even when that means splitting up compound nouns.
Also, both annotators reacted differently to situations where a sequence had to
be marked up. This spawned a new subsection in our annotation manual as well.

@@

Moreover, we came to the conclusion that defining the error context as a set of
tokens might be misleading or at least difficult to understand in case the error
context does not constitute one sequence, but several scattered sequences, such
as proper nouns or syntactic constituents. $C$ could therefore be turned into a
set of sequences of token indices.

Annotating a corpus with this schema is quite laborious, and borders on
impossibility without proper support from an annotation tool, such as the one we
had to devise. During the process, however, our tool was constantly improved
according to the annotators' ideas, and made the annotation process easier in
the process. Partial automation and other features might increase annotation
comfort even further.

>>>>

Upon inspecting the corpus data in greater detail, we found out that it was
sometimes hard to decide whether a particular error was a contextual or
grammatic error. Moreover, the categories within the grammatic error
branch were not clear enough. It seems that our merger of the CLC and Free Text
introduced some confusion with certain categories. Particularly, the
\textit{omission}, \textit{replace}, and \textit{redundancy} cases would at
times `steal' hits from the other categories due to their general
nomenclature. We revised the annotation manual to specifically mention that
all other grammatical error categories take precedence over these three.
Unfortunately, this seems to be a suboptimal solution. The next section contains
a proposal for a much cleaner solution.

@@

However, the error type taxonomy was not our primary concern. In theory, any
taxonomy would fit. We believe that taxonomies with a finer granularity might be
preferable. Our main interest was the agreement on error boundary detection,
error context detection and the three basic error types.  A very unclear point
in that context with respect to the original manual was the minimalization of
token spans constraint, which advises the annotator to only select the smallest
possible amount of tokens, even when that means splitting up compound nouns.
Also, both annotators reacted differently to situations where a sequence had to
be marked up. We discovered that this had quite an effect on the consistency of
our annotations and decided to add a new subsection in our annotation manual
about sequences.

@@

Moreover, we saw that defining the error context as a set of
tokens might be misleading or at least difficult to understand in case the error
context does not constitute one sequence, but several scattered sequences, such
as proper nouns or syntactic constituents. $C$ could therefore be turned into a
set of sequences of token indices.

Annotating a corpus with this schema is quite laborious, and borders on
impossibility without proper support from an annotation tool, such as the one we
had to devise. During the process, however, our tool was constantly improved
according to the annotators' ideas, and made the annotation task easier in
the process. Partial automation and other features might increase annotation
comfort even further.
